Abstract Plots

### 2021-08-19 Abstract.Rmd
Changes in figures and information required for manuscript

#Diagram of Methods

@@

#Diagram of Simulating Data
```{r}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

#Reconstructed Profiles of Signatures 4 and 6

@@

#provide predictions and labels
MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
MMR.perf <- performance(MMR.pred, measure = "tpr", x.measure = "fpr")
tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")

@@

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

```{r}
model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)

#summary(model2)

model2_summary <- summary(model2)

@@

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) |
           signature %in% c("SBS4"))
```

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

@@

#Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

sens.spec.data.2 <- sens.spec.data %>%
  pivot_longer(c(sensitivity, specificity), names_to = "measure", values_to = "value")
```

@@

```{r}
sensitivity = append(sens.MMR.vec, sens.tob.vec)
signature.sens = append(rep("Signature 6", length(sens.MMR.vec)),
                        rep("Signature 4", length(sens.tob.vec)))
dilution = append(step/10, step/10)
sens.df <- data.frame(dilution, sensitivity, signature.sens)

MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Sensitivity as a function of FFPE noise",
       x = "Percent of FFPE noise in samples",
       y = "Sensitivity value",
       color = "True signal") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))

ggsave("MMRtobSens.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
```

@@

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("Signature 6", length(sens.MMR.vec)),
                        rep("Signature 4", length(spec.tob.vec)))
dilution = append(step/10, step/10)
spec.df <- data.frame(dilution, specificity, signature.spec)

MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Specificity as a function of FFPE noise",
       x = "Percent of FFPE noise in samples",
       y = "Specificity value",
       color = "True signal") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))

ggsave("MMRtobSpec.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
```

>>>>

# Diagram of Methods

@@

# Diagram of Simulating Data
```{r}
signature4.plot <- plot_96_profile(signature.4.mat)+
  labs(title = "Signature 4", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

@@

# Reconstructed Profiles of Signatures 4 and 6

@@

#provide predictions and labels
MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
mmr.auc <-  performance(MMR.pred, measure = "auc")
mmr.auc@y.values[[1]][1] #AUC for SBS6
MMR.perf <- performance(MMR.pred, measure = "tpr", x.measure = "fpr")

tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
tob.auc <-  performance(tob.pred, measure = "auc")
tob.auc@y.values[[1]][1] #AUC for SBS4

tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")

@@

# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

```{r}

ggplot(cos.sim.auc.data, aes(x=average_auc)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average AUC", y = "Count")
ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")

median(cos.sim.auc.data$average_auc)
IQR(cos.sim.auc.data$average_auc)

model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)

summary(model2)

@@

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(average_auc<0.8)

cos.sim.auc.outliers

cos.sim.auc.outliers2 <- cos.sim.auc.data %>%
  filter(average_auc>=0.8 & average_auc<0.9)

cos.sim.sbs4 <- cos.sim.auc.data %>%
  filter(signature=="SBS4")

@@

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

# Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE

```{r}
set.seed(500)
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

ggplot(sens.spec.data, aes(x=sensitivity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average sensitivity", y = "Count")
ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")

median(sens.spec.data$sensitivity)
IQR(sens.spec.data$sensitivity)

outliers <- sens.spec.data %>%
  filter(sensitivity<0.8 | signature=="SBS4" | signature=="SBS6")

ref.signatures <- sens.spec.data %>%
  filter(signature=="SBS4" )

ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")

model.sens <- lm(sensitivity ~ cosine_similarity, data = sens.spec.data)

summary(model.sens)

ggplot(sens.spec.data, aes(x=specificity)) +
  geom_histogram(color="black", fill="lightblue")+
  labs( x = "Average specificity", y = "Count")
ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")

median(sens.spec.data$specificity)
IQR(sens.spec.data$specificity)

outliers <- sens.spec.data %>%
  filter(specificity<0.8 | signature=="SBS4" | signature=="SBS6")

ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
  geom_point(data = outliers, color = "red")+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+

ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")

model.spec <- lm(specificity ~ cosine_similarity, data = sens.spec.data)

summary(model.spec)

@@

```{r}
sensitivity = append(sens.MMR.vec, sens.tob.vec)
signature.sens = append(rep("SBS6", length(sens.MMR.vec)),
                        rep("SBS4", length(sens.tob.vec)))
dilution = append(step/10, step/10)
sens.df <- data.frame(dilution, sensitivity, signature.sens)

MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                     y = sensitivity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  labs(x = "Percent of FFPE signature",
       y = "Sensitivity",
       color = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
  scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")

MMR.tob.sens

@@

```{r}
specificity = append(spec.MMR.vec, spec.tob.vec)
signature.spec = append(rep("SBS6", length(sens.MMR.vec)),
                        rep("SBS4", length(spec.tob.vec)))
dilution = append(step/10, step/10)
spec.df <- data.frame(dilution, specificity, signature.spec)

MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                     y = specificity, 
                                     color = signature.sens)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Percent of FFPE signature",
       y = "Specificity",
       color = "Baseline signature") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
    scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")

ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")

MMR.tob.spec
```
